\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{February 6}
\newcommand{\assignmentassignedate}{January 30}
\newcommand{\assignmentnumber}{Two}

\newcommand{\labyear}{2018}
\newcommand{\labday}{Tuesday}
\newcommand{\labtime}{2:30 pm}

\newcommand{\assigneddate}{Assigned: \labday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{SentencesReverser}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/labtwo/SentencesReverser.java}}

\newcommand{\mainprogramhelper}{\lstinline{Sentences}}
\newcommand{\mainprogramhelpersource}{\lstinline{src/main/java/labtwo/Sentences.java}}

% Edit these commands to give the name to the test suite

\newcommand{\testprogram}{\lstinline{TestSentences}}
\newcommand{\testprogramsource}{\lstinline{src/test/java/labtwo/TestSentences.java}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{./gatorgrader.sh --start}}
\newcommand{\gatorgradercheck}{\command{./gatorgrader.sh --check}}

% --> Compiling and running and testing program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradletest}{\command{gradle test}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

% NOTE: Could be improved, problems due to nesting

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/labtwo/SentencesReverser.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 112}
\lhead{Laboratory Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 112\\Introduction to Computer Science II\\
      Spring 2018\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Laboratory \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

To continue to learn how to use GitHub to access the files for an assignment.
You will complete a programming project described in the textbook, following all
of the steps needed to design, implement, and test a sentence-reversing program
in Java. Specifically, you will continue to learn how to implement and test a
Java program and to write a Markdown file, also practicing how to use an
automated grading tool to assess your progress towards correctly completing the
project.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Use the laboratory computers}. The computers in this laboratory feature specialized software for completing
  this course's laboratory and practical assignments. If it is necessary for you to work on a different machine, be sure
  to regularly transfer your work to a laboratory machine so that you can check its correctness. If you cannot use a
  laboratory computer and you need help with the configuration of your own laptop, then please carefully explain its
  setup to a teaching assistant or the course instructor when you are asking questions.

\item {\bf Follow each step carefully}. Slowly read each sentence in the assignment sheet, making sure that you
  precisely follow each instruction. Take notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a teaching assistant or instructor what assignment
  step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the start of a laboratory or practical session
  and then join the appropriate channel. If you have a question about one of the steps in an assignment, then you can
  post it to the designated channel. Or, you can ask a student sitting next to you or talk with a teaching assistant or
  the course instructor.

\item {\bf Store your files in GitHub}. As in previous laboratory assignments, you will be responsible for storing
  all of your files (e.g., Java source code and Markdown-based writing) in a Git repository using GitHub Classroom.
  Please verify that you have saved your source code in your Git repository by using \command{git status} to ensure that
  everything is updated. You can see if your assignment submission meets the established correctness requirements by
  using the provided checking tools on your local computer and in checking the commits in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files, and written reports after you submit them
  through GitHub; you will need them again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Back up your files regularly}. All of your files are regularly backed-up to the servers in the Department of
  Computer Science and, if you commit your files regularly, stored on GitHub. However, you may want to use a flash
  drive, Google Drive, or your favorite backup method to keep an extra copy of your files on reserve. In the event of
  any type of system failure, you are responsible for ensuring that you have access to a recent backup copy of all your
  files.

\item {\bf Explore teamwork and technologies}. While certain aspects of the laboratory assignments will be challenging
  for you, each part is designed to give you the opportunity to learn both fundamental concepts in the field of computer
  science and explore advanced technologies that are commonly employed at a wide variety of companies. To explore and
  develop new ideas, you should regularly communicate with your team and/or the teaching assistants and tutors.

\item {\bf Hone your technical writing skills}. Computer science assignments require to you write technical
  documentation and descriptions of your experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct, remembering that computer scientists must
  effectively communicate and collaborate with their team members and the tutors, teaching assistants, and course
  instructor.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your assignments with others, copying source
  code or writing is a violation of Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

If you have not done so already, please read all of the relevant ``GitHub
Guides'', available at \url{https://guides.github.com/}, that explain how to use
many of the features that GitHub provides. In particular, please make sure that
you have read guides such as ``Mastering Markdown'' and ``Documenting Your
Projects on GitHub''; each of them will help you to understand how to use both
GitHub and GitHub Classroom. To do well on this assignment, you should also read
Chapter 1 in the textbook, paying particularly close attention Sections 1.3 and
1.9. Please see the course instructor or one of the teaching assistants if you
have questions on this reading assignment.

\section*{Accessing the Laboratory Assignment on GitHub}

To access the laboratory assignment, you should go into the
\channel{\#announcements} channel in our Slack team and find the announcement
that provides a link for it. Copy this link and paste it into a web browser.
Now, you should accept the assignment and see that GitHub Classroom created a
new GitHub repository for you to access the assignment's starting materials and
to store the completed version of your assignment. Specifically, to access your
new GitHub repository for this assignment, please click the green ``Accept''
button and then click the link that is prefaced with the label ``Your assignment
has been created here''. If you accepted the assignment and correctly followed
these steps, you should have created a GitHub repository with a name like
``Allegheny-Computer-Science-112-Spring-2018/computer-science-112-spring-2018-lab-2-gkapfham''.
Unless you provide the instructor with documentation of the extenuating
circumstances that you are facing, not accepting the assignment means that you
automatically receive a failing grade for it.

Before you move to the next step of this assignment, please make sure that you
read all of the content on the web site for your new GitHub repository, paying
close attention to the technical details about the commands that you will type
and the output that your program must produce. Now you are ready to download the
starting materials to your laboratory computer. Click the ``Clone or download''
button and, after ensuring that you have selected ``Clone with SSH'', please
copy this command to your clipboard. To enter into your course directory you
should now type \command{cd cs112S2018}. By typing \command{git clone} in your
terminal and then pasting in the string that you copied from the GitHub site you
will download all of the code for this assignment. For instance, if the course
instructor ran the \command{git clone} command in the terminal, it would look
like:

\begin{lstlisting}
  git clone git@github.com:Allegheny-Computer-Science-112-S2018/computer-science-112-spring-2018-lab-2-gkapfham.git
\end{lstlisting}

After this command finishes, you can use \command{cd} to change into the new directory. If you want to \step{go back}
one directory from your current location, then you can type the command \command{cd ..}. Please continue to use the
\command{cd} and \command{ls} commands to explore the files that you automatically downloaded from GitHub. What files
and directories do you see? What do you think is their purpose? Spend some time exploring, sharing your discoveries with
a neighbor and a \mbox{teaching assistant}.

\section*{Implementing and Testing a Sentences Reversing Program}

In order view the source code of a Java program or a Markdown-based writing
assignment, you need a text editor. There are many different text editors on
your workstation and you should feel free to explore these on your own. For
instance, you may adopt either the GVim or the Atom text editor to complete this
laboratory assignment. Please study the source code of the \mainprogramhelper
class, noticing that some of its methods are incomplete. Now, can you draw a
picture showing how these classes (i.e., both test suites and supporting
entities) are related? For instance, you should notice that the
\mainprogramhelper{} has a test suite called \testprogram{} that is written in
the JUnit testing framework; please refer to Section 1.9 and the comments in
the source code of this test suite to learn more about how it works.
Intuitively, a Junit test suite has individual test cases, written as Java
methods, that each work to establish a confidence in the correctness of
\mainprogram{}.

Your program should ultimately produce the output that is given in
Figure~\ref{fig:output}. However, you will note that the current implementation
is missing several necessary methods (e.g., a constructor for the
\program{Sentences} class and then methods like
\program{readSentencesFromFile}). You should add these methods to the program,
including test cases whenever it will help you to establish a confidence in the
correctness of your implementation. If you want to \step{build} your program you
can type the command \gradlebuild{} in your terminal, thereby causing the Java
compiler to check your program for errors and get it ready to run. If you get
any error messages, go back to your text editor and figure out what you
mis-typed and fix it. Once you have solved the problem, make a note of the error
and the solution for resolving it. Re-save your program and then build it again
by re-running the \gradlebuild{}. If you cannot build \mainprogram{} correctly,
then please talk with the course instructor. You can also run the test suite for
the classes in \mainprogram{} by typing \gradletest{} in the terminal. What is
the purpose of each of these test cases? Notice that a test suite does not have
a test for an array of size three---you should add it to ensure correctness!

Once you resolve all of the building and testing errors, you can run your
program by typing \gradlerun{} in the terminal window---this is the ``execute''
step that will run your program and produce the designated output. You can see
if your program is producing the desired output by looking at
Figure~\ref{fig:output}. Do you see that it produces twenty-six lines of output?
Can you see that it performs the array reversal correctly? If not, then repair
the program and re-build and re-run it. Once the program runs, please reflect on
this process. What step did you find to be the most challenging? Why? You should
write your reflections in a file, called \reflection{}, that uses the Markdown
writing language. You can learn more about Markdown by viewing the
aforementioned GitHub guide. To complete this aspect of the assignment, you
should write one high-quality paragraph that reports on your experiences. Your
\reflection{} should also contain answers to other questions about
\mainprogram{}'s implementation and testing.

\begin{figure}[t]
  \centering

  \begin{verbatim}

  Printing Sentences before Reversal ...
    (0, First)
    (1, Second)
    (2, Third)
    (3, Fourth)
    (4, Fifth)
    (5, Sixth)
    (6, Seventh)
    (7, Eighth)
    (8, Ninth)
    (9, Tenth)
  ... Done Printing Sentences before Reversal.
  Reversing the Sentences ...
  ... Done Reversing the Sentences.
  Printing the Sentences after Reversal ...
    (9, Tenth)
    (8, Ninth)
    (7, Eighth)
    (6, Seventh)
    (5, Sixth)
    (4, Fifth)
    (3, Fourth)
    (2, Third)
    (1, Second)
    (0, First)
  ... Done Printing the Sentences after Reversal.

  \end{verbatim}
  \vspace*{-.5in}
  \caption{The Expected Output of the \mainprogram{}.}~\label{fig:output}
  \vspace*{-.25in}
\end{figure}

\section*{Checking the Correctness of Your Program and Writing}

As verified by the Checkstyle tool, the source code for the \mainprogram{} and the
other Java files must adhere to all of the requirements in the Google Java Style
Guide available at \url{https://google.github.io/styleguide/javaguide.html}. The
Markdown file that contains your reflection must adhere to the standards
described in the Markdown Syntax Guide
\url{https://guides.github.com/features/mastering-markdown/}. Finally, your
\reflection{} file should adhere to the Markdown standards established by the
\step{Markdown linting} tool available at
\url{https://github.com/markdownlint/markdownlint/}. Instead of requiring you to
manually check that your deliverables adhere to these industry-accepted
standards, GatorGrader makes it easy for you to automatically check if your
submission meets the standards for correctness. For instance, GatorGrader will
run the provided JUnit tests, check to ensure that \mainprogram{} produces
twenty-six lines of output, and see that you commit to GitHub a sufficient
number of times when completing this assignment.

To get started with the use of GatorGrader, type the command \gatorgraderstart{} in your terminal window. Once this step
completes you can type \gatorgradercheck{}. If your work does not meet all of the assignment's requirements, then you
will see the following output in your terminal: \command{Overall, are there any mistakes in the assignment? Yes}. If you
do have mistakes in your assignment, then you will need to review GatorGrader's output, find the mistake, and try to fix
it. Once your program is building correctly, fulfilling at least some of the assignment's requirements, you should
transfer your files to GitHub using the \gitcommit{} and \gitpush{} commands. For example, if you want to signal that
the \mainprogramsource{} file has been changed and is ready for transfer to GitHub you would first type
\gitcommitmainprogram{} in your terminal, followed by typing \gitpush{} and checking to see that the transfer to GitHub
is successful. If you notice that transferring your code or writing to GitHub did not work correctly, then please try to
determine why, asking a teaching assistant or the course instructor for assistance, if necessary.

After the course instructor enables \step{continuous integration} with a system called Travis CI, when you use the
\gitpush{} command to transfer your source code to your GitHub repository, Travis CI will initialize a \step{build} of
your assignment, checking to see if it meets all of the requirements. If both your source code and writing meet all of
the established requirements, then you will see a green \checkmark{} in the listing of commits in GitHub after awhile.
If your submission does not meet the requirements, a red \naughtmark{} will appear instead. The instructor will reduce a
student's grade for this assignment if the red \naughtmark{} appears on the last commit in GitHub immediately before the
assignment's due date. Yet, if the green \checkmark{} appears on the last commit in your GitHub repository, then you
satisfied all of the main checks, thereby allowing the course instructor to evaluate other aspects of your source code
and writing, as further described in the \step{Evaluation} section of this assignment sheet. Unless you provide the
instructor with documentation of the extenuating circumstances that you are facing, no late work will be considered
towards your grade for this laboratory assignment.

Please note that if you are completing this laboratory assignment on your own laptop the GatorGrader program will only
run correctly if you have installed all of the programs on which it depends. You can see these dependencies (i.e.,
Gradle, Java, Markdownlint, Proselint, and Python) by reading the documentation in your GitHub repository for this
assignment. Note that assignment checking will still work with Travis CI even if you do not have these programs
installed on your laptop. Please see the instructor or teaching assistant if you have trouble installing these programs.

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical writing for any assignment in this course.
Instead, this assignment invites you to submit, using GitHub, the following deliverables.

\begin{enumerate}

  \setlength{\itemsep}{0in}

\item Stored in a Markdown file called \reflection{}, a one-paragraph response,
  each consisting of at least three sentences, for each of the stated questions.

\item A properly documented, well-formatted, and correct version of
  \mainprogramsource{} that both meets all of the established requirements and
  produces the desired output.

\item A properly documented, well-formatted, and correct version all of the
  additional Java files needed to complete this assignment (e.g.,
  \mainprogramhelpersource).

\end{enumerate}

\section*{Evaluation of Your Laboratory Assignment}

Using a report that the instructor shares with you through the commit log in GitHub, you will privately received a grade
on this assignment and feedback on your submitted deliverables. Your grade for the assignment will be a function of the
whether or not it was submitted in a timely fashion and if your program received a green \checkmark{} indicating that it
met all of the requirements. Other factors will also influence your final grade on the assignment. In addition to
studying the efficiency and effectiveness of your Java source code, the instructor will also evaluate the accuracy of
both your technical writing and the comments in your source code. If your submission receives a red \naughtmark{}, the
instructor will reduce your grade for the assignment while still considering the regularity with which you committed to
your GitHub repository and the overall quality of your partially completed work. Please see the instructor if you have
questions about the evaluation of this laboratory assignment.

\section*{Adhering to the Honor Code}

In adherence to the Honor Code, students should complete this assignment on an individual basis. While it is appropriate
for students in this class to have high-level conversations about the assignment, it is necessary to distinguish
carefully between the student who discusses the principles underlying a problem with others and the student who produces
assignments that are identical to, or merely variations on, someone else's work. Deliverables (e.g., Java source code or
Markdown-based technical writing) that are nearly identical to the work of others will be taken as evidence of violating
the \mbox{Honor Code}. Please see the course instructor if you have questions about this policy.

\end{document}
