\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{May 4}
\newcommand{\assignmentassignedate}{April 17}
\newcommand{\assignmentnumber}{Ten}

\newcommand{\labyear}{2018}
\newcommand{\labday}{Tuesday}
\newcommand{\labtime}{2:30 pm}
\newcommand{\labduetime}{11:59 pm}

\newcommand{\assigneddate}{Assigned: \labday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labday, \assignmentduedate, \labyear{} at \labduetime{}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}
\newcommand{\timefile}{\lstinline{writing/time.md}}
\newcommand{\colorfile}{\lstinline{writing/color.md}}
\newcommand{\readme}{\lstinline{README.md}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{./gatorgrader.sh --start}}
\newcommand{\gatorgradercheck}{\command{./gatorgrader.sh --check}}

% --> Compiling and running program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}
\usepackage{fancyvrb}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 112}
\lhead{Final Project}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 112\\Introduction to Computer Science II\\
      Spring 2018\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Final Project \\ \assigneddate{} \\ \duedate{}}

\section*{Introduction}

Throughout the semester, you have explored the fundamentals of algorithm
analysis, data structures, and Java programming by studying, in an applied
fashion, topics such as variables and data types, the creation, use, and testing
of Java classes, conditionals, loops, and recursion and various types of data
containers such as linked lists and trees. This final project invites you to
explore, in greater detail, a real-world application of computer science. You
will learn more about how to use, implement, test, and evaluate different types
of real-world computer software. Since you will complete the final project with
a team, you will also learn more about how GitHub and the Git version control
system can effectively support collaborative software development.

Your project should result in a detailed report that includes all of your source
code, in addition to written materials and technical diagrams that highlight the
key contributions of your work. This technical report should include a
description of why the chosen topic is important and discuss the implementation
and/or experimentation that you undertook. The written material should be
precise, formal, appropriately formatted, grammatically correct, informative,
and interesting. The source code that you write must be carefully documented and
tested. If you install and use existing computer software (e.g., a Java class
for visualizing data structures), the steps for installation and use should be
clearly documented in your report. Also, the report must explain the steps to
run your own Java program. Finally, your report must detail the work completed
by each member of your team; individual contributions should also be reflected
in commits to the team's repository. In addition to writing the aforementioned
final report in Markdown, you will also use Markdown to write and submit a
project proposal and a status update at the intermediate deadlines.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Use the laboratory computers}. The computers in this laboratory feature specialized software for completing
  this course's laboratory and practical assignments. If it is necessary for you to work on a different machine, be sure
  to regularly transfer your work to a laboratory machine so that you can check its correctness. If you cannot use a
  laboratory computer and you need help with the configuration of your own laptop, then please carefully explain its
  setup to a teaching assistant or the course instructor when you are asking questions.

\item {\bf Store your files in GitHub}. As in the past laboratory assignments, you will be responsible for storing all
  of your files (e.g., Java source code and Markdown-based writing) in a Git repository using GitHub Classroom. Please
  verify that you have saved your source code in your Git repository by using \command{git status} to ensure that
  everything is updated. You can see if your assignment submission meets the established correctness requirements by
  using the provided checking tools on your local computer and by checking the commits in GitHub.

\item {\bf Back up your files regularly}. All of your files are regularly backed-up to the servers in the Department of
  Computer Science and, if you commit your files regularly, stored on GitHub. However, you may want to use a flash
  drive, Google Drive, or your favorite backup method to keep an extra copy of your files on reserve. In the event of
  any type of system failure, you are responsible for ensuring that you have access to a recent backup copy of all your
  files.

\item {\bf Explore teamwork and technologies}. While certain aspects of the
  final project will be challenging for you, each part is designed to give you
  the opportunity to learn both fundamental concepts in the field of computer
  science and explore advanced technologies that are commonly employed at a wide
  variety of companies. To explore and develop new ideas, you should regularly
  communicate with your team and/or the teaching assistants and tutors.

\item {\bf Hone your technical writing skills}. Computer science assignments
  require to you write technical documentation and descriptions of your
  experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct,
  remembering that computer scientists must effectively communicate and
  collaborate with their team members and the tutors, teaching assistants, and
  course instructor.

\item {\bf Review all of your past laboratory and practical assignments}. Now
  that you have completed many prior assignments, please review all of your
  prior work to ensure that you understand the concepts needed to explore
  real-world applications of computer science.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your
  assignments with others, copying source code or writing is a violation of
  Allegheny College's Honor Code. Specifically, you should only collaborate on
  this project with the students who are a member of your team. You should not
  share source code or data with students not in your team.

\end{itemize}

\section*{Reading Assignment}

To ensure that you are best prepared to complete this final project, please
review all of the chapters that we have covered up to and after the release date
of this assignment. As we cover new material during the remainder of the
semester (e.g., trees and hashtables), you are also encouraged to review that
content as it will better enable you to complete a high-quality final project.

\section*{Accessing the Final Project Assignment on GitHub}

The students in this class may form their own teams of three or four members;
there should be no teams with two people. Once you pick your team members, you
should designate one individual as the team leader who will complete certain
critical steps that are given later in this assignment sheet. To access the
laboratory assignment, you should first go into the \channel{\#announcements}
channel in our Slack team and find the announcement that provides a link for it.
Now, your team should look into the \channel{\#announcements} channel on Slack
and claim the next available number for your team.

Next, the team leader (i.e., the person you chose to complete this step) will
create their team with the name
\command{Computer-Science-112-Spring-2018-Lab-10-Group-<group label>} and then
accept the project assignment and see that GitHub Classroom created a new
GitHub repository for your team to access the assignment's starting materials
and to store the completed version of your assignment. Note that the team leader
will have to type their group and lab details into a text field. For instance,
if the team leader was in the second group then that person would type ``Group 2
for Lab 10'' into the text field. At this point, each additional member of the
team can accept the assignment through GitHub. Please make sure that each of
your team members joins the correct team---you must be in the team with your
number. Unless you provide the instructor with documentation of the extenuating
circumstances that you are facing, not accepting the assignment means that you
automatically receive a failing grade for it. Please see the instructor with
questions.

Before you move to the next step of this assignment, please make sure that you
read all of the content on the web site for your new GitHub repository, paying
close attention to the technical details about the commands that you will type
and any extra steps that you must take. Now you are ready to download the
starting materials to your laboratory computer. Click the ``Clone or download''
button and, after ensuring that you have selected ``Clone with SSH'', please
copy this command to your clipboard. To enter the correct directory you should
now type \command{cd cs112S2018}. Next, you can type the command \command{ls}
and see that there are some files or directories inside of this directory. By
typing \command{git clone} in your terminal and then pasting in the string that
you copied from the GitHub site you will download all of the code for this
assignment. For the previous example, a student would run a \command{git clone}
command in the terminal in this fashion:

\begin{lstlisting}
  git clone git@github.com:Allegheny-Computer-Science-112-S2018/computer-science-112-spring-2018-lab-10-group-2-for-lab-10.git
\end{lstlisting}

After this command finishes, you can use \command{cd} to change into the new directory. If you want to \step{go back}
one directory from your current location, then you can type the command \command{cd ..}. You may continue to use the
\command{cd} and \command{ls} commands to explore the files that you automatically downloaded from GitHub. Please note
that the course instructor expects students to implement and evaluate all of the Java source code needed to complete
their proposed project. As such, there are no provided Java source code files for this assignment. With that said, you
will need to edit the three Markdown files in the \program{writing/} directory by the stated deadline. Here are some
additional tips to help you design, implement, and test the Java source code for the final project:

\begin{itemize}
  \setlength{\itemsep}{0in}
  \item Make sure that your Java source code contains declarations that place it in the correct package.
  \item Given the name that you picked for your package, ensure that your directories are correct.
  \item Update the \program{build.gradle} file so that it contains the correct package and Java class name.
  \item Update the \program{settings.gradle} file so that it also references the correct directory.
  \item Make sure that your \program{build.gradle} file is setup to support graphical and/or textual output.
  \item Remember that all of your Java source code must meet Google's programming standards.
  \item Recall that there are no GatorGrader checks for this assignment since each project is different.
  \item In GatorGrader's absence, your team should establish correctness checks for your source code.
  \item Importantly, your team must implement JUnit tests to assess
    Java program correctness.
\end{itemize}

\section*{Description of the Topics}

Each team is invited to pick one of the following projects. Please note that a
team selecting the student-designed project must first discuss the idea with the
course instructor, during today's laboratory session, and receive feedback and
then final approval. Please note that your team is fully responsible for
ensuring the feasibility of the final project that it proposes.

% Your team may not pick a final project that one team member previously pursued
% in Computer Science 111.

\begin{enumerate}

  \item {\bf Cryptography and Cryptanalysis}: Explore a topic in the fields that
    make up the ``art and science of sending and decoding secret messages''.
    This project invites your team to implement, test, and evaluate several
    cryptography and/or cryptanalysis systems. To start, you should investigate,
    implement, and test ciphers such as the Caesar and Vigenere ciphers. Then,
    you should use your ciphers to demonstrate that you can successfully send
    secret messages through, for instance, an email server. In addition to
    creating and testing these Java programs, your report should include a
    detailed explanation of how your chosen algorithms work. Finally, you should
    conduct and carefully report on doubling experiments that evaluate the
    efficiency of your tool that performs cryptography and/or cryptanalysis.
    Along with studying content about any other necessary data structures or
    algorithms, students who pick this project should review the material about
    arrays and \program{ArrayList}s in Chapters 3 and 7.

  \item {\bf Data Structures}: This topic invites your team to either
    investigate a data structure that we did not already study this semester or,
    alternatively, further look into a data structure that we previously
    discussed. A team that selects this project might review the textbook's
    content about a specific data structure (e.g., Chapter 7's discussion of
    \program{ArrayList}s) and then identify one area for further study. For
    instance, you could decide to implement and test your own version of this
    data types that is ``backed'' by a fixed-size array. When you implement and
    test your own dynamic array that can both grow and shrink in size, you could
    use as an inspiration some of the guidelines in Chapter 7. Finally, your
    team can run experiments to assess the your implementation's efficiency and
    then write a performance evaluation report.

  \item {\bf Performance Evaluation}: Since it is often important to implement
    computer software that exhibits acceptable time and space overheads, this
    project invites students to use and/or extend performance evaluation
    software. After finding, reading, and understanding textbooks and research
    papers on this topic, students who pick this project should identify a focus
    area and create a benchmarking tool. The final version of the framework
    should allow students to measure the performance of computations in Java
    (perhaps through doubling experiments), report those measurements to the
    benchmark's user, and support informed decision making about design and
    implementation trade-offs. Along with including the source code of the
    benchmarks, this project invites students to write a performance evaluation
    report. Teams picking this project should review Chapter 4's content about
    effective algorithm analysis.

  \item {\bf Searching and Sorting}: In a previous assignment you experimentally
    studied the performance of sorting algorithms. A team that selects this
    project could implement and then analytically (i.e., with a proof) and/or
    empirically study the performance of a sorting algorithm. Or, you could
    investigate an algorithm that searches through a data structure (e.g., a
    \program{Tree} or a \program{hashtable}) and looks for data that matches a
    specific pattern. Students who pick this project should review Section 3.1.2
    and further study Chapter 12 of the textbook. Next, this project invites you
    to conduct doubling experiments to assess the performance of your
    algorithms. Along with including the source code of the doubling experiment
    framework and your chosen algorithm, this project invites students to write
    a performance evaluation report.

  \item {\bf System Implementation}: Now that you understand the key topics of
    this course, you are ready to design, implement, and test your own Java
    program. A team that picks this project should first identify a problem
    domain in which they will work. For instance, you might decide that you want
    to implement your own command-line client for the Twitter social network by
    investigating the Twitter4J library. Then, you will need to decide on the
    data structures and algorithms that you will create to store and manipulate
    your data. Finally, you should conduct experiments to assess the efficiency
    of your implementation. If your team chooses this project, then everyone
    should review Chapter 2 and the chapter(s) for your chosen data structures.
    For instance, if you decide to store Twitter data in a list (e.g., a
    \program{SinglyLinkedList} or an \program{ArrayList}), then you should
    review the relevant details in Chapters 3 and 7.

  \item {\bf Student-Designed Project}: Students will develop an idea for their
    own project that focuses on one or more real-world topics in the field of
    computer science. After receiving the course instructor's approval for your
    idea, you will complete the project and report on your results.

\end{enumerate}

\section*{Final Project Requirements}

To ensure that you and your team have mastered the concepts discussed in this
course, your project's source code should adhere to the following requirements.
These requirements may be modified, at the discretion of the course instructor,
only if a team receives prior permission and documents this approval in the
final report and the source code. Without prior approval, all submitted programs
should contain at least five Java classes that consist in total of at least five
methods in addition to the {\tt main} method. The final project's source code
should include examples of the declaration and use of at least five different
data types (e.g., {\tt int}, {\tt String}, and {\tt boolean}). The program
should also include conditional logic, in the form of {\tt if} or {\tt switch}
statements, and one or more iteration constructs, expressed as {\tt while}, {\tt
do-while}, and {\tt for} loops. Finally, the Java classes must also use either
an array, an {\tt ArrayList}, a {\tt LinkedList}, or some other data structure
that we discussed this semester. The implementation of your final project may
opt to use either a data structure provided by the {\tt java.util} package or
one of those that we implemented and tested during the semester. Additionally,
students must create and provide a means for automatically running at least one
JUnit test suite that contains at minimum ten stand-alone and independent test
cases. Whenever needed, students should perform an analytical algorithm
evaluation and/or conduct and report on doubling experiments that assess the
efficiency of their final project's code.

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical
writing for any assignment in this course. Instead, this assignment invites you
to submit, using GitHub, the following deliverables.

\begin{enumerate}

  \setlength{\itemsep}{0in}

\item Completed, fully commented, and properly formatted versions of all Java
  source code files. Please ensure that you source code adheres to all of the
  requirements mentioned in this assignment sheet (e.g., the use of Java
  constructs and the creation of a JUnit test suite).

\item A one-page written proposal, saved in the file
  \program{writing/proposal.md}, with an informative title, an abstract, a
  description of the main idea, an initial listing of the tasks that you must
  complete, and a plan that you and your team will follow to complete the
  work.

\item A one-paragraph status update, saved in the file
  \program{writing/update.md}, that explains what your team has already
  implemented and the steps that you will take to finish your program.

\item A detailed final project report, saved in the file
  \program{writing/report.md}, that documents, in a project-specific fashion,
  how you designed, implemented, tested, and evaluated your system.

\item Stored in \reflection{}, a four (or five) paragraph reflection on the
  commands that you typed in \command{gvim} or \command{atom} and the terminal
  window. This Markdown-based document should explain the input, output, and
  behavior of each command and the challenges that you confronted when using it.
  For every challenge that you encountered, please explain your solution for it.
  This document should also explain how your team collaborated to finish the
  assignment, with each team member writing their own paragraph inside of this
  file.

\item A commit log in your GitHub repository that clearly shows that the team
  members effectively collaborated. That is, the commit log should show that
  commits were evenly made by all team members; a meaningful photograph of each
  member should appear in the commit log.

\end{enumerate}

\noindent
You must complete all of the aforementioned deliverables by the following due dates:

\begin{enumerate}

  \item {\bf Project Assigned and Project Proposal:} Tuesday, April 17, 2018

    After meeting with the course instructor and team members, pick a topic for
    your final project. Remember, if your team selects the student-designed
    project, you must first have your project approved by the course instructor.
    Next, make sure that you create a GitHub repository that can be accessed by
    the instructor. Finally, write and submit a one-page proposal for your
    project. While you can use the project descriptions on the previous pages as
    a starting point, your proposal should have an informative title, an
    abstract, a description of the main idea, an initial listing of the tasks
    that you must complete, and a plan for completing the work.

  \item {\bf Status Update and Project Demonstration}: Tuesday, April 24, 2018

    As you continue working on your project, please write and submit a one
    paragraph status update through your Git repository. In addition, you should
    give a demonstration, during the laboratory session, highlighting the most
    important code that you have finished.

  \item {\bf Final Project Due Date}: Friday, May 4, 2018 by 11:59 pm

    You should submit the final version of your project through your team's
    GitHub repository. This submission should include all of the relevant source
    code and output, the written reports, and any additional materials that will
    demonstrate the success of your project. While you are encouraged to turn in
    the final project before the examination session starts on the due date,
    students must submit the completed assignment no later than 11:59 pm on the
    due date.

\end{enumerate}

\section*{Evaluation of Your Final Project Assignment}

Using a report that the instructor shares with you through the commit log in GitHub, you will privately received a grade
on this assignment and feedback on your submitted deliverables. Your grade for the assignment will be a function of the
whether or not it was submitted in a timely fashion and if your repository contains correct versions of all of the
required Markdown and source code files. Other factors will also influence your final grade on the assignment. The
instructor will also evaluate the accuracy of both your technical writing and the comments in your source code. Please
see the instructor if you have questions about the evaluation of this final project assignment.

All team members will receive the same baseline grade for the final project. If there are extenuating circumstances in
which one or more of the team members do not effectively collaborate to complete this assignment, then the course
instructor will adjust the grade of specific team members so that it is higher or lower than the baseline grade, as is
fair and necessary. Please see the instructor if you do not understand how he assigns grades for collaborative
assignments. Finally, in adherence to the Honor Code, students should only complete this assignment with their team
members. Deliverables (e.g., Java source code or Markdown-based technical writing) that are nearly identical to the work
of outsiders of your team will be taken as evidence of violating the \mbox{Honor Code}. In summary, please see the
course instructor with any questions about this final project policy.

As previously noted, the instructor will also evaluate the effectiveness with which your team completed this assignment.
The instructor will assess your team's effectiveness as {\bf excellent} when:

\vspace*{-.5em}

\begin{enumerate}
  \setlength{\itemsep}{0pt}

\item All team members contribute correct, useful Java source code and/or Markdown documentation that supports the
  creation of a working and meaningful Java computer program.

\item All team members participate in the laboratory discussions, sharing status updates and ideas.

\item All team members attend the classroom discussions about your current status on the project.

\item All team members regularly communicate with others in a respectful and informative manner.

\item All team members work with others and the instructor to resolve any differences that arise.

\end{enumerate}

\vspace*{-.5em}

The instructor will assess team effectiveness as {\bf good} when the team
members fulfill the aforementioned standards most of the time. The team's
effectiveness will be assessed as {\bf average} when some, but not the majority
of the time, the team members make meaningful contributions to the project,
according to these set standards. The course instructor will assess team
effectiveness as {\bf below average} when, as a whole, the team does not
function according to the above standards.

\end{document}
