\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{February 19}
\newcommand{\assignmentassignedate}{February 16}
\newcommand{\assignmentnumber}{Three}

\newcommand{\labyear}{2018}
\newcommand{\labdueday}{Monday}
\newcommand{\labassignday}{Friday}
\newcommand{\labtime}{1:30 pm}

\newcommand{\assigneddate}{Assigned: \labassignday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labdueday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{BubbleSort}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/practicalthree/BubbleSort.java}}

% Edit these commands to give the name to the test suite

\newcommand{\testprogram}{\lstinline{TestBubbleSort}}
\newcommand{\testprogramsource}{\lstinline{src/test/java/practicalthree/TestBubbleSort.java}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{./gatorgrader.sh --start}}
\newcommand{\gatorgradercheck}{\command{./gatorgrader.sh --check}}

% --> Compiling and running and testing program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradletest}{\command{gradle test}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

% NOTE: Could be improved, problems due to nesting

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/practicalthree/BubbleSort.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 112}
\lhead{Practical Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 112\\Introduction to Computer Science II\\
      Spring 2018\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Practical \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

To continue to practice using GitHub to access the files for a course
assignment. Additionally, to review how to use the Ubuntu operating system and
software development programs such as a ``terminal window'' and the ``text
editor''. You will also continue to practice using Slack to support
communication with the teaching assistants and the instructor. Next, you will
learn how to use a test suite to guide the repair of a Java program that
incorrectly uses nested iteration constructs to sort arrays, also practicing how
to use an automated grading tool to assess your progress.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Use the laboratory computers}. The computers in this laboratory feature specialized software for completing
  this course's laboratory and practical assignments. If it is necessary for you to work on a different machine, be sure
  to regularly transfer your work to a laboratory machine so that you can check its correctness. If you cannot use a
  laboratory computer and you need help with the configuration of your own laptop, then please carefully explain its
  setup to a teaching assistant or the course instructor when you are asking questions.

\item {\bf Follow each step carefully}. Slowly read each sentence in the assignment sheet, making sure that you
  precisely follow each instruction. Take notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a teaching assistant or instructor what assignment
  step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the start of a laboratory or practical session
  and then join the appropriate channel. If you have a question about one of the steps in an assignment, then you can
  post it to the designated channel. Or, you can ask a student sitting next to you or talk with a teaching assistant or
  the course instructor.

\item {\bf Store your files in GitHub}. Starting with this practical assignment, you will be responsible for storing
  all of your files (e.g., Java source code and Markdown-based writing) in a Git repository using GitHub Classroom.
  Please verify that you have saved your source code in your Git repository by using \command{git status} to ensure that
  everything is updated. You can see if your assignment submission meets the established correctness requirements by
  using the provided checking tools on your local computer and in checking the commits in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files, and written reports after you submit them
  through GitHub; you will need them again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Back up your files regularly}. All of your files are regularly backed-up to the servers in the Department of
  Computer Science and, if you commit your files regularly, stored on GitHub. However, you may want to use a flash
  drive, Google Drive, or your favorite backup method to keep an extra copy of your files on reserve. In the event of
  any type of system failure, you are responsible for ensuring that you have access to a recent backup copy of all your
  files.

\item {\bf Explore teamwork and technologies}. While certain aspects of the course assignments will be challenging
  for you, each part is designed to give you the opportunity to learn both fundamental concepts in the field of computer
  science and explore advanced technologies that are commonly employed at a wide variety of companies. To explore and
  develop new ideas, you should regularly communicate with your team and/or the teaching assistants and tutors.

\item {\bf Hone your technical writing skills}. Computer science assignments require to you write technical
  documentation and descriptions of your experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct, remembering that computer scientists must
  effectively communicate and collaborate with their team members and the tutors, teaching assistants, and course
  instructor.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your assignments with others, copying source
  code or writing is a violation of Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

If you have not done so already, please read all of the relevant ``GitHub
Guides'', available at \url{https://guides.github.com/}, that explain how to use
many of the features that GitHub provides. In particular, please make sure that
you have read guides such as ``Mastering Markdown'' and ``Documenting Your
Projects on GitHub''; each of them will help you to understand how to use both
GitHub and GitHub Classroom. To do well on this assignment, you should study the
Wikipedia article about BubbleSort and also read Chapter 3 in the textbook,
paying close attention to Sections 3.1 and 3.6. Please see the course instructor
if you have questions on these reading assignments.

\section*{Accessing the Practical Assignment on GitHub}

To access the laboratory assignment, you should go into the
\channel{\#announcements} channel in our Slack team and find the announcement
that provides a link for it. Copy this link and paste it into a web browser.
Now, you should accept the assignment and see that GitHub Classroom created a
new GitHub repository for you to access the assignment's starting materials and
to store the completed version of your assignment. Specifically, to access your
new GitHub repository for this assignment, please click the green ``Accept''
button and then click the link that is prefaced with the label ``Your
assignment has been created here''. If you accepted the assignment and
correctly followed these steps, you should have created a GitHub repository
with a name like
``Allegheny-Computer-Science-112-Spring-2018/computer-science-112-spring-2018-practical-3-gkapfham''.
Unless you provide the instructor with documentation of the extenuating
circumstances that you are facing, not accepting this practical assignment
means that you automatically receive a failing grade for it.

Before you move to the next step of this assignment, please make sure that you
read all of the content on the web site for your new GitHub repository, paying
close attention to the technical details about the commands that you will type
and the output that your program must produce. Now you are ready to download the
starting materials to your laboratory computer. Click the ``Clone or download''
button and, after ensuring that you have selected ``Clone with SSH'', please
copy this command to your clipboard. At this point, you can open a new terminal
window and type the command \command{cd cs112S2018} to enter the course
directory. Now type \command{git clone} in your terminal and paste in the string
that you copied from the GitHub site to download this assignment's code. For
instance, if the instructor ran the \command{git clone} command, it would look
like:

\begin{lstlisting}
  git clone git@github.com:Allegheny-Computer-Science-112-S2018/computer-science-112-spring-2018-practical-3-gkapfham.git
\end{lstlisting}

After this command finishes, you can use \command{cd} to change into the new
directory. If you want to \step{go back} one directory from your current
location, then you can type the command \command{cd ..}. Please continue to use
the \command{cd} and \command{ls} commands to explore the files that you
automatically downloaded from GitHub. What files and directories do you see?
What do you think is their purpose? Spend some time exploring, sharing your
discoveries with a neighbor and a \mbox{teaching assistant}.

\section*{Repairing a Java Program that Sorts Arrays}

Please study the source code of the \mainprogram{} class. Can you see any
obvious ways in which the implementation of this sorting algorithm is incorrect?
You should run the test suite called \testprogram{} that is written in the JUnit
testing framework. Please read each of the tests as they explain, in Java source
code, the expected behavior of a sorting algorithm. You should pay particularly
close attention to the fact that the test cases use the array-based sorting
algorithm that is provided by the Java programming language to support the
testing of the \mainprogram{}. Why is this a good approach to testing a new
sorting algorithm? You can run the test suite for the \mainprogram{} class by
typing \gradletest{} in the terminal. Did you notice that many of the test cases
do not pass? Can you identify any common reason why the tests might fail? Please
study the source code of \testprogramsource{} to see if you can find and fix any
implementation mistakes in the test suite. Now how many test cases are passing?

At this point, you should notice that many of the tests are still failing. This
means that you will have to examine the source code of \mainprogram{} to
determine what mistakes are lurking in its implementation. Please note that the
two sorting methods contains several (different) mistakes that you should
incrementally fix. Reading about \mainprogram{} on Wikipedia and reasoning about
how the current implementation does not match the algorithm's expected behavior
should help you to find and repair the defects. Remember, if you want to
\step{build} your program you can type the command \gradlebuild{} in your
terminal, thereby causing the Java compiler to check your program for errors and
get it ready to run. If you see any error messages, go back to your text editor
and figure out what you mis-typed and fix it. Once you have solved the problem,
make a note of the error and the solution for resolving it. Re-save your program
and then build it again by re-running the \gradlebuild{}. You should keep
revising the implementation of \mainprogram's methods until all of the tests
pass. Then, you can use \gradlerun{} to see the program's output.

\section*{Checking the Correctness of Your Program and Writing}

As verified by Checkstyle tool, the source code for \mainprogram{} and its test
suite must adhere to all of the requirements in the Google Java Style Guide
available at \url{https://google.github.io/styleguide/javaguide.html}. Instead
of requiring you to manually check that your deliverables adhere to these
industry-accepted standards, GatorGrader makes it easy for you to automatically
check if your submission meets the standards for correctness. For instance,
GatorGrader will run the provided tests, check to ensure that \mainprogram{}
produces exactly four lines of output, and see that you committed to GitHub a
sufficient number of times when completing this assignment.

To get started with the use of GatorGrader, type the command \gatorgraderstart{}
in your terminal window. Once this step completes you can type
\gatorgradercheck{}. If your work does not meet all of the assignment's
requirements, then you will see the following output in your terminal:
\command{Overall, are there any mistakes in the assignment? Yes}. If you do have
mistakes in your assignment, then you will need to review GatorGrader's output,
find the mistake, and try to fix it. Once your program is building correctly,
fulfilling at least some of the assignment's requirements, you should transfer
your files to GitHub using the \gitcommit{} and \gitpush{} commands. For
example, if you want to signal that the \mainprogramsource{} file has been
changed and is ready for transfer to GitHub you would first type
\gitcommitmainprogram{} in your terminal, followed by typing \gitpush{} and
checking to see that the transfer to GitHub is successful. If you notice that
transferring your \mainprogramsource{} to GitHub does not work, then please try
to determine why, asking the course instructor for assistance, if necessary.

After the course instructor enables \step{continuous integration} with a system called Travis CI, when you use the
\gitpush{} command to transfer your source code to your GitHub repository, Travis CI will initialize a \step{build} of
your assignment, checking to see if it meets all of the requirements. If both your source code and writing meet all of
the established requirements, then you will see a green \checkmark{} in the listing of commits in GitHub after awhile.
If your submission does not meet the requirements, a red \naughtmark{} will appear instead. The instructor will reduce a
student's grade for this assignment if the red \naughtmark{} appears on the last commit in GitHub immediately before the
assignment's due date. Yet, if the green \checkmark{} appears on the last commit in your GitHub repository, then you
satisfied all of the main checks, thereby allowing the course instructor to evaluate other aspects of your source code
and writing, as further described in the \step{Evaluation} section of this assignment sheet. Unless you provide the
instructor with documentation of the extenuating circumstances that you are facing, no late work will be considered
towards your grade for this practical assignment.

Please note that if you are completing this practical assignment on your own laptop the GatorGrader program will only
run correctly if you have installed all of the programs on which it depends. You can see these dependencies (i.e.,
Gradle, Java, Markdownlint, Proselint, and Python) by reading the documentation in your GitHub repository for this
assignment. Note that assignment checking will still work with Travis CI even if you do not have these programs
installed on your laptop. Please see the instructor or teaching assistant if you have trouble installing these programs.

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical writing for any assignment in this course.
Instead, this assignment invites you to submit, using GitHub, the following deliverables.

\begin{enumerate}

  \setlength{\itemsep}{0in}

\item A properly documented and correct version of \mainprogramsource{} that
  meets the established requirements and produces the desired textual output.

\item A properly documented and correct version of \testprogramsource{} that
  meets the established requirements and demonstrates sorting method
  correctness.

\end{enumerate}

\section*{Evaluation of Your Practical Assignment}

Using a report that the instructor shares with you through the commit log in
GitHub, you will privately received a grade on this assignment and feedback on
your submitted deliverables. Your grade for the assignment will be a function of
the whether or not it was submitted in a timely fashion and if your program
received a green \checkmark{} indicating that it met all of the requirements.
Other factors will also influence your final grade on the assignment. In
addition to studying the efficiency and effectiveness of your Java source code,
the instructor will also evaluate the accuracy of both your technical writing
and the comments in your source code. If your submission receives a red
\naughtmark{}, the instructor will reduce your grade for the assignment while
still considering the regularity with which you committed to your GitHub
repository and the overall quality of your partially completed work. Please see
the instructor if you have questions about the evaluation of this laboratory
assignment.

In adherence to the Honor Code, students should complete this assignment on an
individual basis. While it is appropriate for students in this class to have
high-level conversations about the assignment, it is necessary to distinguish
carefully between the student who discusses the principles underlying a
technical problem with others and the student who produces assignment
deliverables (e.g., source code) that are identical to, or merely variations on,
someone else's work. 


\end{document}
